# Ch. 12 Linear Regression {-}

Sections covered: 12.1, 12.2, 12.5

## 12.1 The Simple Linear Regression Model {-}

## 12.2 Estimating Model Parameters {-}

Formulas to know from p. 498: 

$b_1 = \dfrac{\sum(x_i -\overline{x})(y_i - \overline{y})}{\sum(x_i - \overline{x})^2} = \frac{S_{xy}}{S_{xx}}$ and $b_0 = \overline{y} - b_1 \overline{x}$

Formula to know from p. 502:

$SSE = \sum(y_i - \hat{y_i})^2$

Formulas to know from p. 504:

$SST = \sum(y_i - \overline{y})^2$ and $r^2 = 1 - \frac{SSE}{SST}$

Formulas to know from p. 505:

$SSR = \sum(\hat{y_i} - \overline{y})^2$ and $SSE + SSR = SST$



### Resources {-}

Interactive Visualization: [Linear Regression](https://jtr13.github.io/D3/BestFittingLine.html) Try fitting the least squares line to a set of random data and check your answer ([and another one](https://www.geogebra.org/m/xC6zq7Zv)).


Video: [Regression I: What is regression? | SSE, SSR, SST | R-squared | Errors (ε vs. e)](https://www.youtube.com/watch?v=aq8VU5KLmkY)


### Textbook p. 507 #17 {-}

Researchers fitted a simple linear regression model to explain how $Y=$ porosity (%) is related to $X=$ unit weight (pcf) in concrete specimens. Consider the following representative data:


```{r}
x <- c(99.0, 101.1, 102.7, 103.0, 105.4, 107.0, 108.7, 110.8, 112.1, 112.4, 113.6, 113.8, 115.1, 115.4, 120.0)
y <- c(28.8, 27.9, 27.0, 25.2, 22.8, 21.5, 20.9, 19.6, 17.1, 18.9, 16.0, 16.7, 13.0, 13.6, 10.8)
```

Using R to find:

*Model coefficients*, $b_0$ and $b_1$:

```{r}
mod <- lm(y ~ x)
mod$coefficients
```

($b_0$ is listed under `(Intercept)` and $b_1$ is listed under `x`.)

*Residuals*

```{r}
mod$residuals
```

rounded:

```{r}
round(mod$residuals, 2)
```

*SSR, SSE, SST*

SSR (regression sum of squares):

```{r}
SSR <- sum((mod$fitted.values - mean(y))^2)
SSR
```


SSE (error sum of squares):

```{r}
SSE <- sum(mod$residuals^2)
SSE
```

SST (total sum of squares):

```{r}
SST <- sum((y - mean(y))^2)
SST
```

(Check that SSR + SSE = SST)

```{r}
SSR + SSE  
```


d. What proportion of observed variation in porosity can be attributed to the approximate linear relationship between unit weight and porosity? 

Method #1: SSR/SST

```{r}
SSR/SST
```

Method #2: $r^2$

```{r}
cor(x, y)^2
```





## 12.5 Correlation {-}

Skip: p. 530 "Inferences About the Population Correlation Coefficient" to the end of the section. 

### Resources {-}

Interactive visualization: [Correlation Coefficient](https://1201.info/ch.-5-joint-probability.html#interactive) (add and remove points)

Interactive visualization: [Interpreting Correlations](https://rpsychologist.com/d3/correlation/) 


### R {-}

Sample correlation coefficient $r$

```{r}
# Example 12.15, p. 528
x <- c(2.4, 3.4, 4.6, 3.7, 2.2, 3.3, 4.0, 2.1)
y <- c(1.33, 2.12, 1.80, 1.65, 2.00, 1.76, 2.11, 1.63)

cor(x,y)
```




