# Ch. 4 Continuous Random Variables and Probability Distributions {-}

Sections covered: 4.1, 4.2, 4.3

## 4.1 Probability density functions {-}

## R {-}

Simulating a uniform distribution

Randomly choose one number from the distribution:

$f(x) = .2$ for $0 \leq x \leq 5$, 0 otherwise.

```{r}
runif(n = 1, min = 0, max = 5)
```

Take the average of 1000 picks:

```{r}
x <- runif(n = 1000, min = 0, max = 5)
mean(x)
```

Draw a histogram of the 1000 picks:
```{r}
hist(x, las = 1, col = "lightblue")
```


## 4.2 Cumulative Distribution Functions and Expected Values {-}

Using R to help with calculations in Example 4.7, p. 149

Define a function for the cdf (you have to do the integration yourself):

```{r}
F <- function(x) (x/8) + (3/16)*x^2
```

Find $F(1.5) - F(1)$

```{r}
F(1.5) - F(1)
```

## 4.3 The Normal Distribution {-}

[Interactive normal distribution](http://www.intmath.com/counting-probability/normal-distribution-graph-interactive.php
) -- change $\mu$ and $\sigma$ and see what happens.

[Normal table](http://users.stat.ufl.edu/~athienit/Tables/Ztable.pdf) -- in pdf form for viewing online, downloading and/or printing

[Online normal distribution lookup](https://www2.ccrb.cuhk.edu.hk/stat/other/normal.htm) -- enter $z$ and it will calculate $F(z) = P(Z \leq z)$.

There are many online normal distribution calculators. Note that often you are given the choice of finding $P(Z \leq z)$, $P(Z \geq z)$, $P(0 \leq Z \leq z)$, etc., so be careful. To be consistent with the normal tables in the textbook / class, choose $P(Z \leq z)$.


Due to rounding differences in $z$ and $P(Z \leq z)$, solutions using technology will vary from those using a normal table. Both methods are fine.

### R {-}

$P(Z \leq -1)$

Unless specified otherwise, `pnorm` uses a mean of 0 and standard deviation of 1 (standard normal).

```{r}
pnorm(-1)
```

$P(X \leq 37)$ given $\mu = 40$ and $\sigma = 2$
```{r}
pnorm(37, mean = 40, sd = 2)
```

or

```{r}
pnorm((37-40)/2)
```

$P(X > 39)$
```{r}
1 - pnorm(39, mean = 40, sd = 2)
```

Find the 75th percentile for the standard normal distribution:
```{r}
qnorm(.75)
```

Find the 75th percentile for a normally distribution population with mean 40 and standard deviation 2:
```{r}
qnorm(.75, mean = 40, sd = 2)
```

or
```{r}
40 + qnorm(.75)*2
```

## Practice Exercises {-}

1. **(pdf, E and V)** $X$ is a continuous random variable with probability distrbution $f(x)=.06x + .05$ where $0\leq x \leq 5$. 

   a. What is the probability that $2\leq x\leq 4$? 
   b. What is the probability that $x = 4$?
   c. What is the expected value of $X$?
   d. What is the variance of X?
   
   
[Ans]
a. Integrate $f(x)=.06x + .05$, we get $F(x) = .03x^2+.05x$.
```{r}
F <- function(x) 0.03*x^2 + 0.05*x
x <- runif(n = 1000, min = 0, max = 5)
F(4) - F(2)
```

b. 0

c. Integrate $x\cdot f(x) = x(.06x+.05)=.06x^2+.05x$, we get $H(x) = .02x^3+.025x^2$.
```{r}
H <- function(x) 0.02*x^3 + 0.025*x^2
H(5)-H(0) 
```

d. $V(X) = E(X^2)-[E(X)]^2$.
Integrate $x^2\cdot f(x) = x^2(.06x+.05)=.06x^3+.05x^2$, we get $G(x) = .015x^4+\frac{.05}3x^3$.
```{r}
E <- H(5)-H(0)
G <- function(x) .015*(x^4)+(.05/3)*(x^3)
G(5)-G(0)-E^2
```



## Extra Practice Chapters 1 - 4 {-}

[Questions](https://github.com/jtr13/1201/blob/master/pdfs/ExtraPractice1-Questions.pdf){target="_blank"}

[Solutions](https://github.com/jtr13/1201/blob/master/pdfs/ExtraPractice1-Solutions.pdf){target="_blank"}
