# Ch. 1 Descriptive Statistics {-} 

Sections covered: all

## 1.1 Populations, Samples, and Processes {-}

## 1.2 Pictorial and Tabular Methods in Descriptive Statistics {-}

Skip: "Dotplots" pp. 15-16

### R {-}

Stem-and-leaf display:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

stem(prices)
```

Frequency histogram:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

hist(prices)

hist(prices, breaks = c(300, 400, 500, 600, 700, 800),
     col = "lightblue")
```

Density histogram:  

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

hist(prices, freq = FALSE, 
     breaks = c(300, 400, 500, 600, 700, 800),
     col = "lightblue", las = 1)

```

## 1.3 Measures of location {-}

### R {-}

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

mean(prices)

median(prices)

## quartiles
quantile(prices)

## trimmed mean
mean(prices, trim = .1) ## 10% trimmed mean
```

## 1.4 Measures of variability {-}

Skip: extreme outliers (p. 42) 

We will define outliers for boxplots to be observations that are more than 1.5 times the fourth spread from the closest fourth. They may be indicated with either a solid or open circle (in contrast to the book which uses one for mild outliers and the other for extreme outliers.)

### R {-}

Sample variance:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

var(prices)
```

Sample standard deviation:

```{r}
sqrt(var(prices))

sd(prices)
```


Five number summary 

(min, lower-hinge, median, upper-hinge, max)

```{r}
fivenum(prices)
```

Boxplots  

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545,
            599, 665, 675, 699, 699, 725, 725, 745, 799)

boxplot(prices)

boxplot(prices, horizontal = TRUE)

PTSD <- c(10, 20, 25, 28, 31, 35, 37, 38, 38, 39, 39, 42, 46)

Healthy <- c(23, 39, 40, 41, 43, 47, 51, 58, 63, 66, 67, 69, 72)

df <- data.frame(Healthy, PTSD)

boxplot(df, horizontal = TRUE)
```

