# Ch. 1 Descriptive Statistics {-} 


```{r, echo = FALSE}
knitr::opts_chunk$set(out.width = '65%', fig.align = 'center')
```

Sections covered: all (skim 1.1)

## 1.2 Pictorial and Tabular Methods in Descriptive Statistics {-}

Skip: Example 1.7, p. 15 (double-digit leaves)

Skip: "Dotplots," pp. 15-16


### Stem-and-leaf plots {-}

**Ex:** The following are the listing prices (in thousands of $) of one-bedroom apartments in Morningside Heights in 2016:
379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665, 675, 699, 699, 725, 725, 745, 799

Draw a stem an leaf plot of this data:

1. Create a new variable (we'll call it "prices"). *Note: we use the syntax <- c() to generate new vectors.*

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)
```

2. Draw a stem and leaf plot

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)
stem(prices)
```

The `scale` parameter will change the number of rows by the factor specified:

```{r}
stem(prices, scale = 2)
stem(prices, scale = .5)
```


### Frequency histogram  {-}

If you want to create a **frequency histogram** of this data you use the following:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

hist(prices)
```

*Note: histograms are drawn with unbinned data. R does the binning in the process of drawing the histogram. This means that the program chooses the size of the bins for you.*


To add specific bin sizes and colors to your histogram, you can use the following syntax:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)
hist(prices, breaks = c(300, 400, 500, 600, 700, 800),
     col = "lightblue")
```

### Density histogram  {-}

We use an almost identical syntax to generate density histograms, but add the parameter `freq=FALSE` -- that is, not a frequency histogram.

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

hist(prices, freq = FALSE, 
     breaks = c(300, 400, 500, 600, 700, 800),
     col = "lightblue", las = 1)

```

(`las = 1` makes all the labels horizontal.)


## 1.3 Measures of location {-}

Skip: Example 1.16, p. 33 (trimmed mean)

Skip: "Categorical Data and Sample Proportions," p. 34 (We'll return to this topic later.)

Consider the same dataset (prices). To find the mean, median, quartiles, and trimmed mean, we use the following syntax:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

mean(prices)

median(prices)

## quartiles
quantile(prices)

## trimmed mean
mean(prices, trim = .1) ## 10% trimmed off each end
```

## 1.4 Measures of variability {-}

Skip: extreme outliers (p. 42) 

We will define outliers for boxplots to be observations that are more than 1.5 times the fourth spread from the closest fourth. They may be indicated with either a solid or open circle (in contrast to the book which uses one for mild outliers and the other for extreme outliers.)

### Numerical summaries  {-}

* Sample variance:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

var(prices)
```

* Sample standard deviation:

```{r}
sqrt(var(prices))

sd(prices)
```

* Five number summary 

(min, lower-hinge, median, upper-hinge, max)

```{r}
fivenum(prices)
```

If you want to generate a horizontal boxplot, use the following 

### Boxplots {-}

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545,
            599, 665, 675, 699, 699, 725, 725, 745, 799)

boxplot(prices)

boxplot(prices, horizontal = TRUE)
```

* Comparitive boxplots

A simple means to create multiple boxplots is to list the vectors that you wish to display. Note that to label the boxplots, use `names = `.


```{r}
PTSD <- c(10, 20, 25, 28, 31, 35, 37, 38, 38, 39, 39, 42, 46)

Healthy <- c(23, 39, 40, 41, 43, 47, 51, 58, 63, 66, 67, 69, 72)

boxplot(PTSD, Healthy, names = c("PTSD", "Healthy"), horizontal = TRUE)
```

## Practice Exercises {-}

1. Using the built-in dataset `ToothGrowth` in R, visualize the data and comment on the effectiveness of different functions in the context.

[Ans] <br/>
```{r}
# The first 5 rows of the data
head(ToothGrowth, 5)

# Five number summary
fivenum(ToothGrowth$len)

# Boxplot 
# '$' extracts the column by name
boxplot(ToothGrowth$len) 

# Stem-and-leaf Plot
stem(ToothGrowth$len)

# Histogram
h <- hist(ToothGrowth$len)

# Cumulative Histogram
h$counts <- cumsum(h$counts)
plot(h)
```

### Summary of R functions {-}

`head()`: directly see how the dataset looks; useful when the dataset is large and it's difficult to display all rows and columns together.

`fivenum()`: returns the minimum value, lower fourth, median, upper fourth, and maximum value

`boxplot()`: visualizes the five number summary plus outliers. (It's clear that the `ToothGrowth` data is not skewed.)

`stem()`: compares the number of data points that fall in different bins. (Here we can see that most values are between 20 and 29.)

`hist()`: draws a histogram -- values are grouped in bins

`cumsum()`: takes a vector and returns the cumulative sums
