# Ch. 4 Continuous Random Variables and Probability Distributions {-}

Sections covered: 4.1, 4.2, 4.3

## 4.1 Probability density functions {-}

## R {-}

Simulating a uniform distribution

Randomly choose one number from the distribution:

$f(x) = .2$ for $0 \leq x \leq 5$, 0 otherwise.

```{r}
runif(n = 1, min = 0, max = 5)
```

Take the average of 1000 picks:

```{r}
x <- runif(n = 1000, min = 0, max = 5)
mean(x)
```

Draw a histogram of the 1000 picks:
```{r}
hist(x, las = 1, col = "lightblue")
```


## 4.2 Cumulative Distribution Functions and Expected Values {-}

Using R to help with calculations in Example 4.7, p. 149

Define a function for the cdf (you have to do the integration yourself):

```{r}
F <- function(x) (x/8) + (3/16)*x^2
```

Find $F(1.5) - F(1)$

```{r}
F(1.5) - F(1)
```

## 4.3 The Normal Distribution {-}

### R {-}

$P(Z \leq -1)$

Unless specified otherwise, `pnorm` uses a mean of 0 and standard deviation of 1 (standard normal).

```{r}
pnorm(-1)
```

$P(X \leq 37)$ given $\mu = 40$ and $\sigma = 2$
```{r}
pnorm(37, mean = 40, sd = 2)
```

or

```{r}
pnorm((37-40)/2)
```

$P(X > 39)$
```{r}
1 - pnorm(39, mean = 40, sd = 2)
```

Find the 75th percentile for the standard normal distribution:
```{r}
qnorm(.75)
```

Find the 75th percentile for a normally distribution population with mean 40 and standard deviation 2:
```{r}
qnorm(.75, mean = 40, sd = 2)
```

or
```{r}
40 + qnorm(.75)*2
```

