# Ch. 3 Discrete Distributions {-}

Sections covered: 3.1 - 3.5 

## 3.3 Expected value {-}

```{r}
x <- 1:5
x
px <- c(.1, .15, .2, .25, .3)
px
x*px
sum(x*px)    # E(X)
```

## 3.3 Variance {-}

```{r}
x - 3.5
(x - 3.5)^2
((x - 3.5)^2)*px
sum(((x - 3.5)^2)*px)   # V(X)
```

## 3.3 Variance (alternative method) {-}
```{r}
x
px
x^2
(x^2)*px
sum((x^2)*px)  # E(X^2)
14-3.5^2     # E(X^2) - [E(X)]^2
```

## 3.4 Binominal Theorem {-}

p. 121 Using binomial tables for the cumulative distribution function (cdf) is optional; you may use R (see below) or [www.stattrek.com](http://www.stattrek.com) instead.

For tests you will not need to calculate these values. You can leave your answers as summations, for example:

$\Sigma_{x=0}^6 \left(\begin{array}{c}12\\ x\end{array}\right)(.3^x)(.7^{12-x})$

### R {-}

Probability mass function (pmf)

$P(X = x)$

```{r}
choose(8, 3)    # "8 choose 3"
56*.6^3*.4^5     # P(X = 3) given n = 8, p = .6
```

Direct method

```{r}
dbinom(3, 8, .6)     # P(X = 3) given n = 8, p = .6
```

Cumulative distribution function (cdf)

$P(X \leq x)$

Find $P(2 \leq X \leq 4)$ given $p = .7, n = 15$

```{r}
pbinom(4, 15, .7) - pbinom(1, 15, .7)
```

(Using the pmf instead:)

```{r}
dbinom(2, 15, .7) + dbinom(3, 15, .7) + dbinom(4, 15, .7)
```

Graphing a binomial pmf:

ex. $p = .7, n = 10$

```{r}
px <- dbinom(0:10, 10, .7)
# adding las = 1 turns the y-axis tick mark labels horizontal, which are easier to read
barplot(px, names.arg = 0:10, las = 1, col = "lightblue")
```



## 3.5 Hypergeometric {-}

Note that the notation that R uses is different from the Devore textbook:

|parameter|Devore|R|
|---------|------|-|
|total successes|M|m|
|total failures|N-M|n|
|sample size|n|k|
|successes in sample|x|x|

Example (p. 127)

Devore: *h(x; n, M, N)*

*P(X = 2) = h(2; 10, 5, 25)* -->

```{r}
dhyper(x = 2, m = 5, n = 20, k = 10)
```

## 3.6 Poisson {-}

Example (p. 132)

p(3;2) =

```{r}
dpois(3,2)
```

F(3;2) =

```{r}
ppois(3, 2)
```

