# Ch. 7 Statistical Intervals Based on a Single Sample {-}

Sections covered: 7.1, 7.2, 7.3

## 7.1 Basic Properties of Confidence Intervals {-}

Skip: "Deriving a Confidence Interval", pp. 282-284; "Bootstrap Confidence Intervals", p. 284

## 7.2 Large-Sample Confidence Intervals for a Population Mean and Proportion {-}

You may use: $\hat{p} \pm z_{\alpha/2} \sqrt{\frac{\hat{p}\hat{q}}{n}}$, rather than the formula (7.10) in the blue box on p. 289 for large sample confidence intervals for proportions, skip p. 290

You may use the method from class $n = \frac{4z^2\hat{p}\hat{q}}{w^2}$ for the minimum $n$ needed to ensure a particular confidence interval width for proportions, rather than formula (7.12) -- both appear on p. 291. As noted in Example 7.9 on p. 291, the easier formula gives a slightly different answer (385 instead of 381).

[Additional textbook material on CI for proportions](https://github.com/jtr13/1201/blob/master/pdfs/LargeSampleCIforPopulationProportion.pdf){target="_blank"}

["When You Hear the Margin of Error Is Plus or Minus 3 Percent, Think 7 Instead"](https://www.nytimes.com/2016/10/06/upshot/when-you-hear-the-margin-of-error-is-plus-or-minus-3-percent-think-7-instead.html){target="_blank"}, New York Times, Oct 5, 2016.

## 7.3 Intervals Based on a Normal Population Distribution {-}

Use the $t$ distributions.

Skip "A Prediction Interval for a Single Future Value" p. 299 to the end of the section

Student's $t$ distribution in R:

The function "dt" will give you the value of the pdf of the $t$ distribution given a certain random variable "X" and degrees of freedom "df". 

Example: Find the value of the $t$ distribution pdf at x = 0 with 20 degrees of freedom:
```{r}
dt(x = 0, df = 20)
```

However, this command will not be as useful as "pt":
The function "pt" will give you the value of the cdf of the $t$ distribution given a certain random variable "X" and degrees of freedom "df". 

Example: Find the area to the left of a t-statistic with value of -0.825 and 14 degrees of freedom.
```{r}
pt(-0.825, 14)
```

Now, if you want to compute the area to the right of a given value X (we'll use the previous example), you can substract the previous command from 1, or add the argument lower.tail = FALSE:

## Method 1
```{r}
1 - pt(-0.825, 14)
```

## Method 2
```{r}
pt(-0.825, 14, lower.tail = FALSE)
```

You might also want to compute the inverse of the cdf of the $t$ distribution. In other words, finding out what the t-score is of the kth quantile of the  $t$ distribution. To do this, you use the "qt" command. The critical values found by this command will match the critical values you see in your t-distribution table.

Example: Find the t-score of the 95th quantile of the $t$ distribution with df = 20
```{r}
qt(.95, df = 20)
```





