# Ch. 1 Descriptive Statistics {-} 

Sections covered: all

## 1.1 Populations, Samples, and Processes {-}

## 1.2 Pictorial and Tabular Methods in Descriptive Statistics {-}

Skip: Example 1.7, p. 15 (double-digit leaves)

Skip: "Dotplots," pp. 15-16

### R {-}

Stem-and-leaf display:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

stem(prices)
```

Note that histograms are drawn with *unbinned* data.  R does the binning in the process of drawing the histogram.

Frequency histogram:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

hist(prices)

hist(prices, breaks = c(300, 400, 500, 600, 700, 800),
     col = "lightblue")
```

Density histogram:  

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

hist(prices, freq = FALSE, 
     breaks = c(300, 400, 500, 600, 700, 800),
     col = "lightblue", las = 1)

```

Cumulative frequency histogram

For this type of histogram, we need access to the bin counts, in order to calculate the cumulative frequencies. The `hist()` function returns these values, if assigned to a variable:

```{r}
x <- c(1, 1, 1, 1, 1, 5, 5, 5, 7, 7, 8)
myhistdata <- hist(x)
myhistdata
```

The particular information we want is `$counts`:

```{r}
myhistdata$counts
```

The cumulative frequencies are:

```{r}
cumsum(myhistdata$counts)
```

To plot them, we need to use a bar chart, not a histogram, since we already have the y-axis values:

```{r}
barplot(cumsum(myhistdata$counts))
```

Cleaned up:

```{r}
barplot(cumsum(myhistdata$counts),
        col = "lightblue", 
        space = 0,  # remove gaps between bars
        las = 1, # make all tick mark labels horizontal
        ylim = c(0, 12), # make the y-axis longer
        names.arg = myhistdata$mids
        )
```



## 1.3 Measures of location {-}

Skip: Example 1.16, p. 33 (trimmed mean)

Skip: "Categorical Data and Sample Proportions," p. 34 (We'll return to this topic later.)

### R {-}

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

mean(prices)

median(prices)

## quartiles
quantile(prices)

## trimmed mean
mean(prices, trim = .1) ## 10% trimmed mean
```

## 1.4 Measures of variability {-}

Skip: extreme outliers (p. 42) 

We will define outliers for boxplots to be observations that are more than 1.5 times the fourth spread from the closest fourth. They may be indicated with either a solid or open circle (in contrast to the book which uses one for mild outliers and the other for extreme outliers.)

### R {-}

Sample variance:

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545, 599, 665,
            675, 699, 699, 725, 725, 745, 799)

var(prices)
```

Sample standard deviation:

```{r}
sqrt(var(prices))

sd(prices)
```


Five number summary 

(min, lower-hinge, median, upper-hinge, max)

```{r}
fivenum(prices)
```

Boxplots  

```{r}
prices <- c(379, 425, 450, 450, 499, 529, 535, 535, 545,
            599, 665, 675, 699, 699, 725, 725, 745, 799)

boxplot(prices)

boxplot(prices, horizontal = TRUE)

PTSD <- c(10, 20, 25, 28, 31, 35, 37, 38, 38, 39, 39, 42, 46)

Healthy <- c(23, 39, 40, 41, 43, 47, 51, 58, 63, 66, 67, 69, 72)

df <- data.frame(Healthy, PTSD)

boxplot(df, horizontal = TRUE)
```

## Practice Exercises {-}

1. Using the built-in dataset `ToothGrowth` in R, visualize the data and comment on the effectiveness of different functions in the context.

```{r}
# The first 5 rows of the data
head(ToothGrowth, 5)

# Five number summary
fivenum(ToothGrowth$len)

# Boxplot 
# '$' extracts the column by name
boxplot(ToothGrowth$len) 

# Stem-and-leaf Plot
stem(ToothGrowth$len)

# Histogram
h <- hist(ToothGrowth$len)

# Cumulative Histogram
h$counts <- cumsum(h$counts)
plot(h)
```

`head()`: directly see how the dataset looks; useful when the dataset is large and it's difficult to display all rows and columns together.

`fivenum()`: returns the minimum value, lower fourth, median, upper fourth, and maximum value

`boxplot()`: visualizes the five number summary plus outliers. (It's clear that the `ToothGrowth` data is not skewed.)

`stem()`: compares the number of data points that fall in different bins. (Here we can see that most values are between 20 and 29.)

`hist()`: draws a histogram -- values are grouped in bins

`cumsum()`: takes a vector and returns the cumulative sums
