# Ch. 4 Continuous Random Variables and Probability Distributions {-}

Sections covered: 4.1, 4.2, 4.3

## 4.1 Probability density functions {-}

## R {-}

Simulating a uniform distribution

Randomly choose one number from the distribution:

$f(x) = .2$ for $0 \leq x \leq 5$, 0 otherwise.

```{r}
runif(n = 1, min = 0, max = 5)
```

To take 1000 picks (to give you an estimate of the mean) from a distribution with the previous pdf:

```{r}
x <- runif(n = 1000, min = 0, max = 5)
mean(x)
```

To draw a histogram of the 1000 picks:
```{r}
hist(x, las = 1, col = "lightblue")
```
_Note: this histogram won't show a perfect uniform distribution because it would need more picks (infinity) to perfectly resemble a histogram._

If you wanted to calculate the uniform probability density function in an interval (1 to 3 in this example) for a sample of values (numbers from 0 to 4 in this example) you can use the following:
```{r}
x <- 0:4 
dunif(x, min = 1, max = 3)
```
_Note: as seen in the output, for those numbers that aren't in the interval (0 and 4), the resulting probabilities were 0. For the rest, they were 0.5 _

**Cumulative uniform function:**
In R, you can calculate the uniform cumulative distribution function:

Example:
 X is the time (in minutes) that a person has to wait in order to take a flight and it is uniformly distributed in the interval (0,60). Find the probability of waiting less than 15 minutes for a flight:
```{r}
punif(15, min = 0, max = 60)
```

**Finding the quantile in a uniform distributions**
You can calculate the corresponding quantile for any probability for a uniform distribution.

Example:
Consider the previous uniform distribution of the flight waiting times. In case you want to calculate the 20% quantile for this distribution (20% of the waiting times will be below this value) you can type:
```{r}
qunif(0.2, min = 0, max = 60)
```

To find the median (50% quantile) using this function you can type:
```{r}
qunif(0.5, min = 0, max = 60)
```

## 4.2 Cumulative Distribution Functions and Expected Values {-}

Using R to help with calculations in Example 4.7, p. 149

Define a function for the cdf (you have to do the integration yourself):

```{r}
F <- function(x) (x/8) + (3/16)*x^2
```

Find $F(1.5) - F(1)$

```{r}
F(1.5) - F(1)
```

## 4.3 The Normal Distribution {-}

[Interactive normal distribution](http://www.intmath.com/counting-probability/normal-distribution-graph-interactive.php
) -- change $\mu$ and $\sigma$ and see what happens.

[Normal table](http://users.stat.ufl.edu/~athienit/Tables/Ztable.pdf) -- in pdf form for viewing online, downloading and/or printing

[Online normal distribution lookup](https://www2.ccrb.cuhk.edu.hk/stat/other/normal.htm) -- enter $z$ and it will calculate $F(z) = P(Z \leq z)$.

There are many online normal distribution calculators. Note that often you are given the choice of finding $P(Z \leq z)$, $P(Z \geq z)$, $P(0 \leq Z \leq z)$, etc., so be careful. To be consistent with the normal tables in the textbook / class, choose $P(Z \leq z)$.


Due to rounding differences in $z$ and $P(Z \leq z)$, solutions using technology will vary from those using a normal table. Both methods are fine.

### R {-}

$P(Z \leq -1)$

Unless specified otherwise, `pnorm` uses a mean of 0 and standard deviation of 1 (standard normal).

```{r}
pnorm(-1)
```

$P(X \leq 37)$ given $\mu = 40$ and $\sigma = 2$
```{r}
pnorm(37, mean = 40, sd = 2)
```

or

```{r}
pnorm((37-40)/2)
```

$P(X > 39)$
```{r}
1 - pnorm(39, mean = 40, sd = 2)
```

Find the 75th percentile for the standard normal distribution:
```{r}
qnorm(.75)
```

Find the 75th percentile for a normally distribution population with mean 40 and standard deviation 2:
```{r}
qnorm(.75, mean = 40, sd = 2)
```

or
```{r}
40 + qnorm(.75)*2
```

## Practice Exercises {-}


1. **(pdf, E and V)** $X$ is a continuous random variable with probability distrbution $f(x)=.06x + .05$ where $0\leq x \leq 5$. 

   a. What is the probability that $2\leq x\leq 4$? 
   b. What is the probability that $x = 4$?
   c. What is the expected value of $X$?
   d. What is the variance of X?
   
[Ans]
a. Integrating by hand: $f(x)=.06x + .05$, we get $F(x) = .03x^2+.05x$.

```{r}
F <- function(x) 0.03*x^2 + 0.05*x
x <- runif(n = 1000, min = 0, max = 5)
F(4) - F(2)
```

b. 0

c. Integrate $x\cdot f(x) = x(.06x+.05)=.06x^2+.05x$, we get $H(x) = .02x^3+.025x^2$.
```{r}
H <- function(x) 0.02*x^3 + 0.025*x^2
H(5)-H(0) 
```

d. $V(X) = E(X^2)-[E(X)]^2$.
Integrate $x^2\cdot f(x) = x^2(.06x+.05)=.06x^3+.05x^2$, we get $G(x) = .015x^4+\frac{.05}3x^3$.
```{r}
E <- H(5)-H(0)
G <- function(x) .015*(x^4)+(.05/3)*(x^3)
G(5)-G(0)-E^2
```
------

2. **(Standard Normal, `pnorm`)** A driver's reaction time for an in-traffic response to a brake signal can be modeled with a normal distribution having mean value 1.25 sec and standard deviation of .46 sec.  (Textbook 4.16)

   a. What is the probability that reaction time is between 1.00 sec and 1.75 sec?

   b. What is the probability that reaction time exceeds 2.00 sec?

```{r}
pnorm((1.75-1.25)/0.46) - pnorm((1-1.25)/0.46)
1 - pnorm((2-1.25)/0.46)
# alternatively
pnorm(-(2-1.25)/0.46)
```
------

3. **(Nonstandard Normal, `qnorm`)** Data collected from an experiment with a specified initial crack
   length propose a normal distribution with mean value 5.496 mm and standard deviation .067 mm. For this model, what value of final crack depth would be exceeded by only .5% of all cracks? (Textbook 4.18)

```{r}
qnorm(0.995, 5.496, .067)
```
------

4. **(Binom Appproximation)** Suppose that 25% of all students at a large public university receive financial aid. Let $X$ be the number of students in a random sample of size 50 who receive financial
   aid, sothat *p = .25*. What is the probability that at most 10 students receive aid? (Textbook 4.20)

```{r}
# binom distribution calculation
pbinom(10, size = 50, prob = 0.25) 
# normal approximation of binom
pnorm((10 + .5 - 50*.25)/sqrt(50*.25*.75))

```

## Extra Practice Chapters 1 - 4 {-}

[Questions](https://github.com/jtr13/1201/blob/master/pdfs/ExtraPractice1-Questions.pdf){target="_blank"}

[Solutions](https://github.com/jtr13/1201/blob/master/pdfs/ExtraPractice1-Solutions.pdf){target="_blank"}
