# Ch. 12 Linear Regression {-}

Sections covered: 12.1, 12.2, 12.5

## 12.1 The Simple Linear Regression Model {-}

## 12.2 Estimating Model Parameters {-}

Formulas to know from p. 498: 

$b_1 = \dfrac{\sum(x_i -\overline{x})(y_i - \overline{y})}{\sum(x_i - \overline{x})^2} = \frac{S_xy}{S_xx}$ and $b_0 = \overline{y} - b_1 \overline{x}$

Formula to know from p. 502:

$SSE = \sum(y_i - \hat{y_i})^2$

Formulas to know from p. 504:

$SST = \sum(y_i - \overline{y})^2$ and $r^2 = 1 - \frac{SSE}{SST}$

Formulas to know from p. 505:

$SSR = \sum(\hat{y_i} - \overline{y})^2$ and $SSE + SSR = SST$



### Resources {-}

Interactive Visualization: [Linear Regression](https://jtr13.github.io/D3/BestFittingLine.html) Try fitting the least squares line to a set of random data and check your answer ([and another one](https://www.geogebra.org/m/xC6zq7Zv)).


Video: [Regression I: What is regression? | SSE, SSR, SST | R-squared | Errors (ε vs. e)](https://www.youtube.com/watch?v=aq8VU5KLmkY) [contributed by Lance J.]

### R {-}

Calculating slope and intercept for a sample of (x, y) pairs (p. 498 formulas)

```{r}
# Example 12.8, p. 503
x <- c(12, 30, 36, 40, 45, 57, 62, 67, 71, 78, 93, 94, 100, 105)
y <- c(3.3, 3.2, 3.4, 3, 2.8, 2.9, 2.7, 2.6, 2.5, 2.6, 2.2, 2, 2.3, 2.1)
lm(y ~ x)  #lm = linear model
```

**Predicted values:**

```{r}
mod <- lm(y~x)
round(mod$fitted.values, 2)
```

**Residuals:**

```{r}
round(mod$residuals, 2)
```

**SSE, SSR**

```{r}
anova(mod)  # anova = analysis of variance
```
The first row under "Sum Sq" is the SSR, and the second row under "Sum Sq" is the SSE:

SSE = `r anova(mod)[2,2]`

SSR = `r anova(mod)[1,2]`

SST = SSE + SSR = `r anova(mod)[2,2]` + 
`r anova(mod)[1,2]` = `r round(anova(mod)[2,2] + anova(mod)[1,2], 4)`

**Coefficient of determination $r^2$**

```{r}
# Example 12.4, 12.9
x <- c(132, 129, 120, 113.2, 105, 92, 84, 83.2, 88.4, 59, 80, 81.5, 71, 69.2)
y <- c(46, 48, 51, 52.1, 54, 52, 59, 58.7, 61.6, 64, 61.4, 54.6, 58.8, 58)

mod <- lm(y ~ x)

SSR <- anova(mod)$`Sum Sq`[1]
SST <- anova(mod)$`Sum Sq`[1] + anova(mod)$`Sum Sq`[2]
SSR/SST
```

Or (simply):
```{r}
cor(x,y)^2
```

(See section 12.5)


## 12.5 Correlation {-}

Skip: "Inferences About the Population Correlation Coefficient" (p. 530) *to end of section*.

### Resources {-}

Interactive visualization: [Correlation Coefficient](https://jtr13.github.io/D3/CorrelationCoefficient.html) (add and remove points)

Interactive visualization: [Interpreting Correlations](https://rpsychologist.com/d3/correlation/) [contributed by Dario G.]


### R {-}

Sample correlation coefficient $r$

```{r}
# Example 12.15, p. 528
x <- c(2.4, 3.4, 4.6, 3.7, 2.2, 3.3, 4.0, 2.1)
y <- c(1.33, 2.12, 1.80, 1.65, 2.00, 1.76, 2.11, 1.63)

cor(x,y)
```


## Practice Exercises {-}
1. **(Least squares line)** Researchers employed a least squares analysis in studying how $Y=$ porosity (%) is related to $X=$ unit weight (pcf) in concrete specimens. Consider the following representative data:


```{r}
x <- c(99.0, 101.1, 102.7, 103.0, 105.4, 107.0, 108.7, 110.8, 112.1, 112.4, 113.6, 113.8, 115.1, 115.4, 120.0)
y <- c(28.8, 27.9, 27.0, 25.2, 22.8, 21.5, 20.9, 19.6, 17.1, 18.9, 16.0, 16.7, 13.0, 13.6, 10.8)
```
(Textbook 12.17)

a. Obtain the equation of the estimated regression line. 

```{r}
lm(y~x)
```
$y = 118.91 - 0.9047x$  

b. Calculate the residuals corresponding to the first two observations.


```{r}
mod <- lm(y~x)
round(mod$residuals, 2)
```

Or alternatively, use R as a calculator
```{r}
pred <- 118.9099 - 0.9047*x
res <- y - pred
res[1]
res[2]
```

c. Calculate a point estimate of $\sigma$.

```{r}
sig2 <- sum((res)^2)/(length(x)-2)
sqrt(sig2)
```

d. What proportion of observed variation in porosity can be attributed to the approximate linear relationship between unit weight and porosity? 

```{r}
cor(x, y)^2
```

e. Calculate the SSE and SST.

```{r}
anova(mod) # analsis of variance
SSE <- anova(mod)$`Sum Sq`[1]
SST <- anova(mod)$`Sum Sq`[1] + anova(mod)$`Sum Sq`[2]
c(SSE, SST)
```

Or alternatively, use R as a calculator. Notice that the same results are produced.
```{r}
SSE1 <- sum((mod$residual)^2)
SST1 <- sum((y-mean(y))^2)
SSR1 <- sum((mod$fitted.values - mean(y))^2)
c(SSE1, SST1, SSE1+SSR1)
```


