# Ch. 12 Linear Regression {-}

Sections covered: 12.1, 12.2, 12.5

## 12.1 The Simple Linear Regression Model {-}

## 12.2 Estimating Model Parameters {-}

### Resources {-}

Interactive Visualization: [Linear Regression](https://www.geogebra.org/m/xC6zq7Zv) Try fitting the least squares line to a set of random data and check your answer.

Video: [Regression I: What is regression? | SSE, SSR, SST | R-squared | Errors (ε vs. e)](https://www.youtube.com/watch?v=aq8VU5KLmkY) [contributed by Lance J.]

### R {-}

Calculating slope and intercept for a sample of (x, y) pairs (p. 498 formulas)

```{r}
# Example 12.8, p. 503
x <- c(12, 30, 36, 40, 45, 57, 62, 67, 71, 78, 93, 94, 100, 105)
y <- c(3.3, 3.2, 3.4, 3, 2.8, 2.9, 2.7, 2.6, 2.5, 2.6, 2.2, 2, 2.3, 2.1)
lm(y ~ x)  #lm = linear model
```

Predicted values:

```{r}
mod <- lm(y~x)
round(mod$fitted.values, 2)
```

Residuals:

```{r}
round(mod$residuals, 2)
```

SSE, SSR

```{r}
anova(mod)  # anova = analysis of variance
```


SSE = `r anova(mod)[2,2]`

SSR = `r anova(mod)[1,2]`

SST = SSE + SSR = `r anova(mod)[2,2]` + 
`r anova(mod)[1,2]` = `r round(anova(mod)[2,2] + anova(mod)[1,2], 4)`

coefficient of determination $r^2$

```{r}
# Example 12.4, 12.9
x <- c(132, 129, 120, 113.2, 105, 92, 84, 83.2, 88.4, 59, 80, 81.5, 71, 69.2)
y <- c(46, 48, 51, 52.1, 54, 52, 59, 58.7, 61.6, 64, 61.4, 54.6, 58.8, 58)

mod <- lm(y ~ x)

SSE <- anova(mod)[2,2]
SST <- anova(mod)[1,2] + anova(mod)[2,2]
1 - (SSE/SST)
```

Or (simply):
```{r}
cor(x,y)^2
```

(See section 12.5)


## 12.5 Correlation {-}

Skip: "Inferences About the Population Correlation Coefficient" (p. 530) *to end of section*.

### Resources {-}

Interactive visualization: [Correlation Coefficient](https://jtr13.github.io/D3/CorrelationCoefficient.html) (add and remove points)

Interactive visualization: [Interpreting Correlations](https://rpsychologist.com/d3/correlation/) [contributed by Dario G.]


### R {-}

sample correlation coefficient $r$

```{r}
# Example 12.15, p. 528
x <- c(2.4, 3.4, 4.6, 3.7, 2.2, 3.3, 4.0, 2.1)
y <- c(1.33, 2.12, 1.80, 1.65, 2.00, 1.76, 2.11, 1.63)

cor(x,y)
```

